---
title: "HW3"
author: "Sijia Yue"
date: "11/5/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(e1071)
library(ada)
```

```{r}
df_train = read_table("synth.tr.txt") %>% 
  mutate(yc = as.factor(yc)) %>% 
  as.data.frame()
df_test = read_table("synth.te.txt") %>% 
  mutate(yc = as.factor(yc)) %>% 
  as.data.frame()
```

## Construct a linear support vector classifier

```{r}
set.seed(313)
# tune the lowest cost
tune_cost = tune(svm, yc ~., data = df_train, kernel = "linear", range = list(cost = c(1:10)))
tune_cost
# train the model
svm_Linear <- svm(yc~., data = df_train, kernel = "linear", scale = FALSE, cost = 2)
# predict on test set
test_pred <- predict(svm_Linear, df_test)
# get the test error
test_error = mean(test_pred != df_test$yc)
# get the standard error of the test error
se = sd(test_pred != df_test$yc)/sqrt(nrow(df_test))
# visualize the predction
plot(svm_Linear, df_test)
```

The test error is `r round(test_error, 3)` and the standard error is `r round(se, 4)`.

## Construct a support vector classifier with Radial kernel.

```{r}
set.seed(333)
# tune the lowest cost
tune_cost = tune(svm, yc ~., data = df_train, kernel = "radial", range = list(cost = c(1:10)))
# train the model
svm_radial <- svm(yc~., data = df_train, kernel = "radial", scale = FALSE, cost = 8)
# predict on test set
test_pred <- predict(svm_radial, df_test)
# get the test error
test_error = mean(test_pred != df_test$yc)
# get the standard error of the test error
se = sd(test_pred != df_test$yc)/sqrt(nrow(df_test))
# visualize the predction
plot(svm_radial, df_test)
```

The test error is `r round(test_error, 3)` and the standard error is `r round(se, 4)`.

## Construct a classifier using AdaBoost algorithm (with 50 boosting iterations) with decision stumps as weak learners.

```{r}
set.seed(333)
# train the model
ada_fit <- ada(yc ~., data = df_train, iter = 50)
# predict on test set
test_pred <- predict(ada_fit, df_test)
# get the test error
test_error = mean(test_pred != df_test$yc)
# get the standard error of the test error
se = sd(test_pred != df_test$yc)/sqrt(nrow(df_test))
# visualize the predction
plot(ada_fit)
```

The test error is `r round(test_error, 3)` and the standard error is `r round(se, 4)`.
